---
title: "Machine Learning Project"
author: "Michal Heydel"
date: "29 November 2017"
output: pdf_document
---
ignore this libaries, after i need to pick the one we use
```{r warning=FALSE, message=FALSE}
require(Matrix) # matrix transformations
require(glmnet) # ridge, lasso & elastinet
require(xgboost) # gbm
require(randomForest)
require(Metrics) # rmse
require(dplyr) # load this in last so plyr doens't overlap it
require(caret) # one hot encoding
require(scales) # plotting $$
require(e1071) # skewness
require(corrplot) # correlation plot
```

```{r}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

```

```{r}
setwd("C:/Users/Michal/Documents/01- Master Degree/GitHub/ST443-Project-group9/Housing price data")
getwd()
```
```{r}
train = read.csv("train.csv", row.names = "Id", stringsAsFactors=FALSE)
testing_kaggle = read.csv("test.csv", row.names = "Id", stringsAsFactors=FALSE)
```

```{r}
#combining train and test data for quicker data prep
testing_kaggle$SalePrice <- NA
train$isTrain <- 1
testing_kaggle$isTrain <- 0
df <- rbind(train,testing_kaggle)
```

```{r}
hist(df$SalePrice)
```

```{r}
colSums(sapply(df, is.na))
```

```{r warning=FALSE, message=FALSE}
df[,c('PoolQC','PoolArea')] %>%
  group_by(PoolQC) %>%
  summarise(mean = mean(PoolArea), counts = n()) 
```
```{r}
df[is.na(df$Exterior1st),]
```


```{r PoolArea}
df[(df$PoolArea > 0) & is.na(df$PoolQC),c('PoolQC','PoolArea')]
```
```{r}
df[2421,'PoolQC'] = 'Ex'
df[2504,'PoolQC'] = 'Ex'
df[2600,'PoolQC'] = 'Fa'
df$PoolQC[is.na(df$PoolQC)] = 'None'
```


```{r}
garage.cols <- c('GarageArea', 'GarageCars', 'GarageQual', 'GarageFinish', 'GarageCond', 'GarageType')
#df[is.na(df$GarageCond),garage.cols]
```
*Garage Yrblt is NAs when there is no garage in the house - set it to?
```{r GarageYrBlt}
#length(which(df$GarageYrBlt == df$YearBuilt))
df[(df$GarageArea > 0) & is.na(df$GarageYrBlt), c(garage.cols, 'GarageYrBlt')]
df$GarageYrBlt[2127] <- df$YearBuilt[2127]
df[2127, 'GarageQual'] <- Mode(df$GarageQual)
df[2127, 'GarageFinish'] <- Mode(df$GarageFinish)
df[2127, 'GarageCond'] <- Mode(df$GarageCond)
df$GarageYrBlt[which(is.na(df$GarageYrBlt))] <- 0
```


to numeric - 0, to categorical = 'None'
```{r}
for(i in garage.cols){
if (sapply(df[i], is.numeric) == TRUE){
    df[,i][which(is.na(df[,i]))] <- 0
  }
  else{
    df[,i][which(is.na(df[,i]))] <- "None"
  }
}
```

```{r KitchenQual}
df$KitchenQual[which(is.na(df$KitchenQual))] <- Mode(df$KitchenQual)
```

```{r MSZoning MSSubClass}
df[is.na(df$MSZoning),c('MSZoning','MSSubClass')]
table(df$MSZoning, df$MSSubClass)
```
```{r}
df$MSZoning[c(2217, 2905)] = 'RL'
df$MSZoning[c(1916, 2251)] = 'RM'
```


There are 486 Nas in LotFrontage, cannot do much - set the NAs to median. Maybe one can do some grouping.

```{r LotFrontage}
df$LotFrontage[which(is.na(df$LotFrontage))] <- median(df$LotFrontage,na.rm = T)
```

there are 2721  NAs in Alley cannot do much - set them equal to "None"
```{r Alley}
df$Alley[which(is.na(df$Alley))] <- "None"
```

one is missing the rest as before
```{r MasVnrArea MasVnrType}
#df[(df$MasVnrArea > 0) & (is.na(df$MasVnrType)),c('MasVnrArea','MasVnrType')]
df[2611, 'MasVnrType'] = 'BrkFace'
df$MasVnrType[is.na(df$MasVnrType)] = 'None'
df$MasVnrArea[is.na(df$MasVnrArea)] = 0
```

for small number of NAs we apply Mode to the categorical, and median to the continous
```{r}
for(i in colnames(df[,sapply(df, is.character)])){
  if (sum(is.na(df[,i])) < 5){
    df[,i][which(is.na(df[,i]))] <- Mode(df[,i])
  }
}
```


```{r}
for(i in colnames(df[,sapply(df, is.integer)])){
  if (sum(is.na(df[,i])) < 5){
    df[,i][which(is.na(df[,i]))] <- median(df[,i], na.rm = T)
  }
}
```

for large number of NAs we apply string "None" to the categorical as a seperate Level, and 0 to the continous

```{r}
for(i in colnames(df[,sapply(df, is.character)])){
    df[,i][which(is.na(df[,i]))] <- "None"
}
```

we have filled in all the missing values. The remaining ones are the SalesPrice in the predicting Dataset <- is fine!
```{r}
colSums(sapply(df, is.na))
sum(is.na(df)) == 1459
```
#### THE DATA IMPUTING IS DONE BY NOW. END------------------------------------------------------------------

###### changing ordinal ones ---------------------------------------------------------------
```{r}
str(df)
```
```{r}
train <- df[df$isTrain==1,]
group.prices <- function(col) {
  group.table <- train[,c(col, 'SalePrice', 'OverallQual')] %>%
    group_by_(col) %>%
    summarise(mean.Quality = round(mean(OverallQual),2),
      mean.Price = mean(SalePrice), n = n()) %>%
    arrange(mean.Quality)
    
  print(qplot(x=reorder(group.table[[col]], -group.table[['mean.Price']]), y=group.table[['mean.Price']]) +
    geom_bar(stat='identity', fill='cornflowerblue') +
    theme_minimal() +
    scale_y_continuous(labels = dollar) +
    labs(x=col, y='Mean SalePrice') +
    theme(axis.text.x = element_text(angle = 45)))
  
  return(data.frame(group.table))
}
```

```{r}
# function that maps a categoric value to its corresponding numeric value and returns that column to the data frame
map.fcn <- function(cols, map.list, dataframe1){
  for (col in cols){
    dataframe1[col] <- as.numeric(map.list[df[,col]])
  }
  return(dataframe1)
}
```

```{r BsmtExposure}
bsmt_list <- c('None' = 0, 'No' = 1, 'Mn' = 2, 'Av' = 3, 'Gd' = 4)

df = map.fcn(c('BsmtExposure'), bsmt_list, df)
```

```{r}
bsmt_fin_list <- c('None' = 0, 'Unf' = 1, 'LwQ' = 2,'Rec'= 3, 'BLQ' = 4, 'ALQ' = 5, 'GLQ' = 6)
df <- map.fcn(c('BsmtFinType1','BsmtFinType2'), bsmt_fin_list, df)
```

```{r QUALITY COLUMNS}
qual.cols <- c('ExterQual', 'ExterCond', 'GarageQual', 'GarageCond', 'FireplaceQu', 'KitchenQual', 'HeatingQC', 'BsmtQual')

qual.list <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)

df<- map.fcn(qual.cols, qual.list, df)
```

```{r warning=FALSE, message=FALSE}
group.prices('Functional')
functional.list <- c('None' = 0, 'Sal' = 1, 'Sev' = 2, 'Maj2' = 3, 'Maj1' = 4, 'Mod' = 5, 'Min2' = 6, 'Min1' = 7, 'Typ'= 8)

df$Functional <- as.numeric(functional.list[df$Functional])
```
```{r warning=FALSE, message=FALSE}
group.prices('GarageFinish')
garage.fin.list <- c('None' = 0,'Unf' = 1, 'RFn' = 1, 'Fin' = 2)

df$GarageFinish <- as.numeric(garage.fin.list[df$GarageFinish])


group.prices('Fence')
fence.list <- c('None' = 0, 'MnWw' = 1, 'GdWo' = 1, 'MnPrv' = 2, 'GdPrv' = 4)

df$Fence <- as.numeric(fence.list[df$Fence])

group.prices('MSSubClass')

MSdwelling.list <- c('20' = 3, '30'= 0, '40' = 1, '45' = 0,'50' = 1, '60' = 4, '70' = 2, '75' = 3, '80' = 2, '85' = 1, '90' = 1, '120' = 3, '150' = 0, '160' = 1, '180' = 0, '190' = 1)

df$MSSubClass <- as.numeric(MSdwelling.list[as.character(df$MSSubClass)])
```
#We've transformed all the categoric features with an ordinal scale into numeric columns. 
####---------------------------------------------------------------------------------------------------



```{r}
for(i in colnames(df[,sapply(df, is.character)])){
    df[,i] <- as.factor(df[,i])
}
```


```{r}
#df$MSSubClass <- as.factor(df$MSSubClass)
#df$OverallQual <- as.factor(df$OverallQual)
#df$OverallCond <- as.factor(df$OverallCond)

```


```{r}
str(df)
```

```{r, eval = FALSE}
Column_classes <- sapply(names(df),function(x){class(df[[x]])})
numeric_columns <-names(Column_classes[Column_classes != "factor"])

#determining skew of each numric variable
skew <- sapply(numeric_columns,function(x){skewness(df[[x]],na.rm = T)})

# Let us determine a threshold skewness and transform all variables above the treshold.
skew <- skew[skew > 0.75]

# transform excessively skewed features with log(x + 1)
for(x in names(skew)) {
  df[[x]] <- log(df[[x]] + 1)
}
```
```{r}
skew
```



```{r}
str(df)
```


### DON't USE THIS CODE -> EVAL = FALSE
##### THIS CODE IS BRILLIANT IT TURN ALL THE FACTOR TO NUMERICAL VALUES FROM 1 to max(factor_level)
```{r, eval = FALSE}
df <- data.frame(lapply(df, function(x) as.numeric(x)))
```



