---
title: "Data Cleaning"
output: html_document
---

```{r}
library(data.table)
library(Matrix)
library(dplyr)
```

```{r}
train = read.csv("train.csv", row.names = "Id", stringsAsFactors=FALSE)
testing_kaggle = read.csv("test.csv", row.names = "Id", stringsAsFactors=FALSE)
#combining train and test data for quicker data prep
testing_kaggle$SalePrice <- NA
train$isTrain <- 1
testing_kaggle$isTrain <- 0
df <- rbind(train,testing_kaggle)
```


```{r}
hist(df$SalePrice)
```

```{r}
colSums(sapply(df, is.na))
```

```{r}
for(i in colnames(df[,sapply(df, is.character)])){
    
  df[,i][which(is.na(df[,i]))] <- "None"
}
```


```{r}
colSums(sapply(df, is.na))
```


```{r}
df$LotFrontage[which(is.na(df$LotFrontage))] <- median(df$LotFrontage,na.rm = T)
df$MasVnrArea[which(is.na(df$MasVnrArea))] <- mean(df$LotFrontage,na.rm = T)
x = c("BsmtFinSF1","BsmtFinSF2", "BsmtUnfSF", "TotalBsmtSF", "BsmtFullBath", "BsmtHalfBath", "GarageYrBlt", "GarageCars", "GarageArea")

for(i in x){
    
  df[,i][which(is.na(df[,i]))] <- 0
}
```


```{r}
colSums(sapply(df, is.na))
```

#Data Cleaning
MSzoning Clustering

```{r}
price <- summarize(group_by(train, train$MSZoning),
          mean(SalePrice, na.rm=T))

price[order(price$`mean(SalePrice, na.rm = T)`),]

df$zone[df$MSZoning %in% c("FV")] <- "Floating Village"
df$zone[df$MSZoning %in% c("RL")] <- "Residential Low Density"
df$zone[df$MSZoning %in% c("RH","RM")] <- "Residential Medium and High Density"
df$zone[df$MSZoning %in% c("C (all)")] <- "Others"
```

Street

```{r}
df$paved[df$Street == "Pave"] <- "Y"
df$paved[df$Street != "Pave"] <- "N"
```

Alley: Binary

```{r}
price <- summarize(group_by(train, Alley),
                   mean(SalePrice, na.rm=T))

price[order(price$`mean(SalePrice, na.rm = T)`),]

df$alleypave[df$Alley %in% c("Pave")] <- "Y"
df$alleypave[!df$Alley %in% c("Pave")] <- "N"
```

LotShape - Combined all the irregular together

```{r}
df$regshape[df$LotShape == "Reg"] <- "Y"
df$regshape[df$LotShape != "Reg"] <- "N"
```

LandContour - Classified it Flat vs Non Flat

```{r}
df$flat[df$LandContour == "Lvl"] <- "Y"
df$flat[df$LandContour != "Lvl"] <- "N"
```

Utilities - All Utilities vs  Missing

```{r}
df$pubutil[df$Utilities == "AllPub"] <- "Y"
df$pubutil[df$Utilities != "AllPub"] <- "N"
```

LandSlope - Gentle or Not

```{r}
df$gentle_slope[df$LandSlope == "Gtl"] <- "Y"
df$gentle_slope[df$LandSlope != "Gtl"] <- "N"
```

No Idea what he did here

```{r groupmeans_lotconfig, echo=FALSE}
summarize(group_by(train, LotConfig),
           mean(SalePrice, na.rm=T))

df$culdesac_fr3[df$LandSlope %in% c("CulDSac", "FR3")] <- "CUl-de-sac or Frontage on 3 sides of property"
df$culdesac_fr3[!df$LandSlope %in% c("CulDSac", "FR3")] <- "Inside/Corner Lot or Frontage on 2 sides of property"
```

No idea what he did here either.

```{r groupmeans_nbhd, echo=FALSE}
nbhdprice <- summarize(group_by(train, Neighborhood),
         mean(SalePrice, na.rm=T))

nbhdprice[order(nbhdprice$`mean(SalePrice, na.rm = T)`),]

nbhdprice_lo <- filter(nbhdprice, nbhdprice$`mean(SalePrice, na.rm = T)` < 140000)
nbhdprice_med <- filter(nbhdprice, nbhdprice$`mean(SalePrice, na.rm = T)` < 200000 &
                          nbhdprice$`mean(SalePrice, na.rm = T)` >= 140000 )
nbhdprice_hi <- filter(nbhdprice, nbhdprice$`mean(SalePrice, na.rm = T)` >= 200000)

df$nbhd_price_level[df$Neighborhood %in% nbhdprice_lo$Neighborhood] <- "Low"
df$nbhd_price_level[df$Neighborhood %in% nbhdprice_med$Neighborhood] <- "Medium"
df$nbhd_price_level[df$Neighborhood %in% nbhdprice_hi$Neighborhood] <- "High"

summarize(group_by(train, Condition1),
           mean(SalePrice, na.rm=T))

df$pos_features_1[df$Condition1 %in% c("PosA", "PosN")] <- "Good"
df$pos_features_1[!df$Condition1 %in% c("PosA", "PosN")] <- "Bad"

summarize(group_by(train, Condition2),
           mean(SalePrice, na.rm=T))

df$pos_features_2[df$Condition1 %in% c("PosA", "PosN")] <- "Good"
df$pos_features_2[!df$Condition1 %in% c("PosA", "PosN")] <- "Bad"

```

Combined single-family detached and townhouse end unit, null otherwise

```{r groupmeans_bldg, echo=FALSE}
summarize(group_by(train, BldgType),
           mean(SalePrice, na.rm=T))

df$twnhs_end_or_1fam[df$BldgType %in% c("1Fam", "TwnhsE")] <- "Y"
df$twnhs_end_or_1fam[!df$BldgType %in% c("1Fam", "TwnhsE")] <- "N"
```

Divide housestyle into 3 bands from its continuous value

```{r}
housestyle_price <- summarize(group_by(train, HouseStyle),
         mean(SalePrice, na.rm=T))

housestyle_lo <- filter(housestyle_price, housestyle_price$`mean(SalePrice, na.rm = T)` < 140000)
housestyle_med <- filter(housestyle_price, housestyle_price$`mean(SalePrice, na.rm = T)` < 200000 &
                          housestyle_price$`mean(SalePrice, na.rm = T)` >= 140000 )
housestyle_hi <- filter(housestyle_price, housestyle_price$`mean(SalePrice, na.rm = T)` >= 200000)

df$house_style_level[df$HouseStyle %in% housestyle_lo$HouseStyle] <- "Low"
df$house_style_level[df$HouseStyle %in% housestyle_med$HouseStyle] <- "Medium"
df$house_style_level[df$HouseStyle %in% housestyle_hi$HouseStyle] <- "High"
```

Change roofstyle into a binary variable based off the mean price.
Hip and shed 1, else 0

```{r}
roofstyle_price <- summarize(group_by(train, RoofStyle),
         mean(SalePrice, na.rm=T))

df$roof_hip_shed[df$RoofStyle %in% c("Hip", "Shed")] <- "Y"
df$roof_hip_shed[!df$RoofStyle %in% c("Hip", "Shed")] <- "N"
```

Changed roof material into binary.
Membrane, Wood Shakes Wood Shingles 1
Else 0

```{r}
roofmatl_price <- summarize(group_by(train, RoofMatl),
         mean(SalePrice, na.rm=T))

df$roof_matl_hi[df$RoofMatl %in% c("Membran", "WdShake", "WdShngl")] <- "Membrane, Wood Shakes Wood Shingles"
df$roof_matl_hi[!df$RoofMatl %in% c("Membran", "WdShake", "WdShngl")] <- "Others"
```

Regrouped the different categories of exterior covering into 3 bands: Dont understand his syntax

```{r}
price <- summarize(group_by(train, Exterior1st),
         mean(SalePrice, na.rm=T))

matl_lo_1 <- filter(price, price$`mean(SalePrice, na.rm = T)` < 140000)
matl_med_1<- filter(price, price$`mean(SalePrice, na.rm = T)` < 200000 &
                          price$`mean(SalePrice, na.rm = T)` >= 140000 )
matl_hi_1 <- filter(price, price$`mean(SalePrice, na.rm = T)` >= 200000)

df$exterior_1[df$Exterior1st %in% matl_lo_1$Exterior1st] <- "Low"
df$exterior_1[df$Exterior1st %in% matl_med_1$Exterior1st] <- "Med"
df$exterior_1[df$Exterior1st %in% matl_hi_1$Exterior1st] <- "High"
```

Regrouped the different categories of exterior if more than one material into 3 bands: 

```{r}
price <- summarize(group_by(train, Exterior2nd),
          mean(SalePrice, na.rm=T))

matl_lo <- filter(price, price$`mean(SalePrice, na.rm = T)` < 140000)
matl_med <- filter(price, price$`mean(SalePrice, na.rm = T)` < 200000 &
                          price$`mean(SalePrice, na.rm = T)` >= 140000 )
matl_hi <- filter(price, price$`mean(SalePrice, na.rm = T)` >= 200000)

df$exterior_2[df$Exterior2nd %in% matl_lo$Exterior2nd] <- "Low"
df$exterior_2[df$Exterior2nd %in% matl_med$Exterior2nd] <- "Med"
df$exterior_2[df$Exterior2nd %in% matl_hi$Exterior2nd] <- "High"
```

Masonry Veneer Type into a binary variable
Stone and Brick Face = 1
Else = 0

```{r}
price <- summarize(group_by(train, MasVnrType),
          mean(SalePrice, na.rm=T))

df$exterior_mason_1[df$MasVnrType %in% c("Stone", "BrkFace") | is.na(df$MasVnrType)] <- "Stone and Brick Face"
df$exterior_mason_1[!df$MasVnrType %in% c("Stone", "BrkFace") & !is.na(df$MasVnrType)] <- "Others"
```

Renamed the external quality of the material categories into 5 numerical levels. NA=0, No Poor observations

```{r}
price <- summarize(group_by(train, ExterQual),
          mean(SalePrice, na.rm=T))

df$exterior_cond[df$ExterQual == "Ex"] <- "Excellent"
df$exterior_cond[df$ExterQual == "Gd"] <- "Good"
df$exterior_cond[df$ExterQual == "TA"] <- "Typical"
df$exterior_cond[df$ExterQual == "Fa"] <- "Fair"
```

Relabeled the external condition of the material categories into 5 numerical levels. NA=0

```{r}
price <- summarize(group_by(train, ExterCond),
          mean(SalePrice, na.rm=T))

df$exterior_cond2[df$ExterCond == "Ex"] <- "Excellent"
df$exterior_cond2[df$ExterCond == "Gd"] <- "Good"
df$exterior_cond2[df$ExterCond == "TA"] <- "Typical" 
df$exterior_cond2[df$ExterCond == "Fa"] <- "Fair"
df$exterior_cond2[df$ExterCond == "Po"] <- "Poor"
```

Changed type of foundation into a binary variable
Pconc = 1

```{r}

price <- summarize(group_by(train, Foundation),
          mean(SalePrice, na.rm=T))

df$found_concrete[df$Foundation == "PConc"] <- "Poured Concrete"
df$found_concrete[df$Foundation != "PConc"] <- "Others"
```

Basement Height: Setting poor as default.

```{r}
price <- summarize(group_by(train, BsmtQual),
          mean(SalePrice, na.rm=T))

df$bsmt_cond1[df$BsmtQual == "Ex"] <- "Excellent"
df$bsmt_cond1[df$BsmtQual == "Gd"] <- "Good"
df$bsmt_cond1[df$BsmtQual == "TA"] <- "Typical"
df$bsmt_cond1[df$BsmtQual == "Fa"] <- "Fair"
df$bsmt_cond1[is.na(df$BsmtQual)] <- "No Basement"
```

Basement Condition: Setting Excellent as default

```{r}
summarize(group_by(train, BsmtCond),
          mean(SalePrice, na.rm=T))

df$bsmt_cond2[df$BsmtCond == "Gd"] <- 5
df$bsmt_cond2[df$BsmtCond == "TA"] <- 4
df$bsmt_cond2[df$BsmtCond == "Fa"] <- 3
df$bsmt_cond2[is.na(df$BsmtCond)] <- 2
df$bsmt_cond2[df$BsmtCond == "Po"] <- 1
```

Basment Exposure: 

```{r}
price <- summarize(group_by(train, BsmtExposure),
          mean(SalePrice, na.rm=T))

df$bsmt_exp[df$BsmtExposure == "Gd"] <- 5
df$bsmt_exp[df$BsmtExposure == "Av"] <- 4
df$bsmt_exp[df$BsmtExposure == "Mn"] <- 3
df$bsmt_exp[df$BsmtExposure == "No"] <- 2
df$bsmt_exp[is.na(df$BsmtExposure)] <- 1
```

Basement rating1:

```{r}
price <- summarize(group_by(train, BsmtFinType1),
          mean(SalePrice, na.rm=T))

df$bsmt_fin1[df$BsmtFinType1 == "GLQ"] <- 5
df$bsmt_fin1[df$BsmtFinType1 == "Unf"] <- 4
df$bsmt_fin1[df$BsmtFinType1 == "ALQ"] <- 3
df$bsmt_fin1[df$BsmtFinType1 %in% c("BLQ", "Rec", "LwQ")] <- 2
df$bsmt_fin1[is.na(df$BsmtFinType1)] <- 1
```

Basement rating2:

```{r}
price <- summarize(group_by(train, BsmtFinType2),
          mean(SalePrice, na.rm=T))

df$bsmt_fin2[df$BsmtFinType2 == "ALQ"] <- 6
df$bsmt_fin2[df$BsmtFinType2 == "Unf"] <- 5
df$bsmt_fin2[df$BsmtFinType2 == "GLQ"] <- 4
df$bsmt_fin2[df$BsmtFinType2 %in% c("Rec", "LwQ")] <- 3
df$bsmt_fin2[df$BsmtFinType2 == "BLQ"] <- 2
df$bsmt_fin2[is.na(df$BsmtFinType2)] <- 1
```

Changed type of heating into a binary

```{r hvac, echo=FALSE}
price <- summarize(group_by(train, Heating),
          mean(SalePrice, na.rm=T))


df$gasheat[df$Heating %in% c("GasA", "GasW")] <- "Y"
df$gasheat[!df$Heating %in% c("GasA", "GasW")] <- "N"
```

Heating Quality

```{r}
price <- summarize(group_by(train, HeatingQC),
          mean(SalePrice, na.rm=T))

df$heatqual[df$HeatingQC == "Ex"] <- 5
df$heatqual[df$HeatingQC == "Gd"] <- 4
df$heatqual[df$HeatingQC == "TA"] <- 3
df$heatqual[df$HeatingQC == "Fa"] <- 2
df$heatqual[df$HeatingQC == "Po"] <- 1
```

Central Air Conditioning: Binary

```{r}
df$air[df$CentralAir == "Y"] <- "Y"
df$air[df$CentralAir == "N"] <- "N"
```

Electrical System

```{r}
price <- summarize(group_by(train, Electrical),
          mean(SalePrice, na.rm=T))

df$standard_electric[df$Electrical == "SBrkr" | is.na(df$Electrical)] <- "Circuit Breaker"
df$standard_electric[!df$Electrical == "SBrkr" & !is.na(df$Electrical)] <- "Others"
```

Kitchen Quality

```{r}
price <- summarize(group_by(train, KitchenQual),
          mean(SalePrice, na.rm=T))

df$kitchen[df$KitchenQual == "Ex"] <- 4
df$kitchen[df$KitchenQual == "Gd"] <- 3
df$kitchen[df$KitchenQual == "TA"] <- 2
df$kitchen[df$KitchenQual == "Fa"] <- 1
```

Fireplace quality

```{r}
price <- summarize(group_by(train, FireplaceQu),
          mean(SalePrice, na.rm=T))

df$fire[df$FireplaceQu == "Ex"] <- 5
df$fire[df$FireplaceQu == "Gd"] <- 4
df$fire[df$FireplaceQu == "TA"] <- 3
df$fire[df$FireplaceQu == "Fa"] <- 2
df$fire[df$FireplaceQu == "Po" | is.na(df$FireplaceQu)] <- 1
```

Garage Type: Binary

```{r}

price <- summarize(group_by(train, GarageType),
          mean(SalePrice, na.rm=T))

df$gar_attach[df$GarageType %in% c("Attchd", "BuiltIn")] <- "Attached or Built in"
df$gar_attach[!df$GarageType %in% c("Attchd", "BuiltIn")] <- "Others"
```

Garage interior: Binary

```{r}
price <- summarize(group_by(train, GarageFinish),
          mean(SalePrice, na.rm=T))

df$gar_finish[df$GarageFinish %in% c("Fin", "RFn")] <- "Y"
df$gar_finish[!df$GarageFinish %in% c("Fin", "RFn")] <- "N"
```

Garage Quality: 5 bands

```{r}
price <- summarize(group_by(train, GarageQual),
          mean(SalePrice, na.rm=T))

df$garqual[df$GarageQual == "Ex"] <- 5
df$garqual[df$GarageQual == "Gd"] <- 4
df$garqual[df$GarageQual == "TA"] <- 3
df$garqual[df$GarageQual == "Fa"] <- 2
df$garqual[df$GarageQual == "Po" | is.na(df$GarageQual)] <- 1
```

Garage Condition: 5 bands

```{r}
price <- summarize(group_by(train, GarageCond),
          mean(SalePrice, na.rm=T))

df$garqual2[df$GarageCond == "Ex"] <- 5
df$garqual2[df$GarageCond == "Gd"] <- 4
df$garqual2[df$GarageCond == "TA"] <- 3
df$garqual2[df$GarageCond == "Fa"] <- 2
df$garqual2[df$GarageCond == "Po" | is.na(df$GarageCond)] <- 1
```

Paved driveway: Binary Paved vs Unpaved

```{r}
price <- summarize(group_by(train, PavedDrive),
          mean(SalePrice, na.rm=T))

df$paved_drive[df$PavedDrive == "Y"] <- "Y"
df$paved_drive[!df$PavedDrive != "Y"] <- "N"
df$paved_drive[is.na(df$paved_drive)] <- "N"
```

Home Functionality: Binary

```{r}
price <- summarize(group_by(train, Functional),
          mean(SalePrice, na.rm=T))

df$housefunction[df$Functional %in% c("Typ", "Mod")] <- 1
df$housefunction[!df$Functional %in% c("Typ", "Mod")] <- 0
```

Pool Quality: Binary 

```{r}
price <- summarize(group_by(train, PoolQC),
          mean(SalePrice, na.rm=T))

df$pool_good[df$PoolQC %in% c("Ex")] <- "Y"
df$pool_good[!df$PoolQC %in% c("Ex")] <- "N"
```

Fence Quality: Binary

```{r}
price <- summarize(group_by(train, Fence),
          mean(SalePrice, na.rm=T))

df$priv_fence[df$Fence %in% c("GdPrv")] <- "Y"
df$priv_fence[!df$Fence %in% c("GdPrv")] <- "N"
```

Sale Type: 5 bands

```{r}
price <- summarize(group_by(train, SaleType),
          mean(SalePrice, na.rm=T))

price[order(price$`mean(SalePrice, na.rm = T)`),]

df$sale_cat[df$SaleType %in% c("New", "Con")] <- 5
df$sale_cat[df$SaleType %in% c("CWD", "ConLI")] <- 4
df$sale_cat[df$SaleType %in% c("WD")] <- 3
df$sale_cat[df$SaleType %in% c("COD", "ConLw", "ConLD")] <- 2
df$sale_cat[df$SaleType %in% c("Oth")] <- 1
```

Sale Condition: 4 bands

```{r}
price <- summarize(group_by(train, SaleCondition),
         mean(SalePrice, na.rm=T))

price[order(price$`mean(SalePrice, na.rm = T)`),]

df$sale_cond[df$SaleCondition %in% c("Partial")] <- 4
df$sale_cond[df$SaleCondition %in% c("Normal", "Alloca")] <- 3
df$sale_cond[df$SaleCondition %in% c("Family","Abnorml")] <- 2
df$sale_cond[df$SaleCondition %in% c("AdjLand")] <- 1
```


Drop all the old variables

```{r}

df$Street <- NULL
df$LotShape <- NULL
df$LandContour <- NULL
df$Utilities <- NULL
df$LotConfig <- NULL
df$LandSlope <- NULL
df$Neighborhood <- NULL
df$Condition1 <- NULL
df$Condition2 <- NULL
df$BldgType <- NULL
df$HouseStyle <- NULL
df$RoofStyle <- NULL
df$RoofMatl <- NULL

df$Exterior1st <- NULL
df$Exterior2nd <- NULL
df$MasVnrType <- NULL
df$ExterQual <- NULL
df$ExterCond <- NULL

df$Foundation <- NULL
df$BsmtQual <- NULL
df$BsmtCond <- NULL
df$BsmtExposure <- NULL
df$BsmtFinType1 <- NULL
df$BsmtFinType2 <- NULL

df$Heating <- NULL
df$HeatingQC <- NULL
df$CentralAir <- NULL
df$Electrical <- NULL
df$KitchenQual <- NULL
df$FireplaceQu <- NULL

df$GarageType <- NULL
df$GarageFinish <- NULL
df$GarageQual <- NULL
df$GarageCond <- NULL
df$PavedDrive <- NULL

df$Functional <- NULL
df$PoolQC <- NULL
df$Fence <- NULL
df$MiscFeature <- NULL
df$SaleType <- NULL
df$SaleCondition <- NULL
df$MSZoning <- NULL
df$Alley <- NULL
```

```{r}
for(i in colnames(df[,sapply(df, is.character)])){
    df[,i] <- as.factor(df[,i])
}
# These are also categorical Variables
df$MSSubClass <- as.factor(df$MSSubClass)
df$OverallCond <- as.factor(df$OverallCond)
df$OverallQual <- as.factor(df$OverallQual)
str(df)
```